---
title: "netlit  Vignette"
author: "Devin Judge-Lord, Adeline Lo & Kyler Hudson"
subtitle: Redistricting Literature
output:
  # pdf_document:
  #   toc: yes
  html_document:
    highlight: zenburn
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(visNetwork)
library(igraph)
library(googlesheets4)
library(fontawesome)

# update data from google sheet if token if  present

source(here::here("code", "update.R"))

# The R package files are currently in NetLit/Netlit on dropbox

library(netlit)

knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE, #FIXME when updating data 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='../figs/',
                      fig.retina = 1, #FIXME change to 8 for paper-quality figs (BUT HIGHER RESOLUTION IS TOO BIG FOR GIT)
                      warning=FALSE, 
                      message=FALSE)


# Table formatting
library(kableExtra)
kablebox <- . %>% 
  slice_head(n = 100) %>%
  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```


# Network Analysis of the Redistricting Literature Data

```{r data}
# load data
literature <- here("Data",  "literature.csv") %>% 
  read_csv() %>% 
  filter(cite_weight > 0)

node_attributes <- read_csv(here("data", "node_attributes.csv"))
```


There are multiple packages to work with networks, but the most popular one used is often `igraph` because of its flexibility and ease in usage Other commonly used packages include `sna` and `networks`. This package has dependencies from the `igraph` package and the user will need to have an up-to-date installation of it.

To use igraph functions, we first need to create an `igraph` object. We can use the `graph_from_data_frame` function, which takes two arguments: `d`, a data frame with the edge list in the first two columns; and `vertices`, a data frame with node data with the node label in the first column. (igraph calls the nodes `vertices`, but it's exactly the same thing.)

```{r}
g <- graph_from_data_frame(d = literature %>% 
                        dplyr::select(from, 
                                      to, 
                                      everything())#, vertices = node_attributes
                      )

g
 
# attributes of the nodes
vertex_attr(g) %>% 
  as_tibble() %>% 
  head()
```

From this `igraph` object, one can use `igraph` functions to generate network statistics. However, this involves several steps and the output is not tidy. To simplify generating network statistics from literature reviews, we introduce the `netlit` package. 

# Using the `netlit` R Package 


The `netlit`  package provides functions to generate network statistics from a literature review. Specifically, `netlit` provides a wrapper for `igraph` functions to facilitate using network analysis in literature reviews. 


Install this package with 
```
devtools::install_github("judgelord/netlit")
```

To install `netlit` from CRAN, run the following:

```{r, eval=FALSE}
install.packages("netlit")
```

## Basic Usage

The `review()` function takes in a dataframe, `data`, that includes `from` and `to` columns (a directed graph structure). 

In the example below, we use example data from [this project on redistricting](https://github.com/judgelord/redistricting). These data are a set of related concepts (`from` and `to`) in the redistricting literature and citations for these relationships (`cites` and `cites_empirical`). See `vignette("netlit")` for more details on this example.

```{r}
library(netlit)

data("literature")

head(literature)
```

`netlit` offers four main functions: `make_edgelist()`, `make_nodelist()`, `augment_nodelist()`, and `review()`. 

`review()` is the primary function. The others are helper functions that perform the individual steps that `review()` does all at once. `review()` takes in a dataframe with at least two columns representing linked concepts (e.g., a cause and an effect) and returns data augmented with network statistics. Users must either specify "from" nodes and "to" nodes with the `from` and `to` arguments or include columns named `from` and `to` in the supplied `data` object.

`review()` returns a list of three objects: 

1. an augmented `edgelist` (a list of relationships with `edge_betweenness` calculated), 
2. an augmented `nodelist` (a list of concepts with `degree` and `betweenness` calculated), and 
3. a `graph` object suitable for use in other `igraph` functions or other network visualization packages. 

Users may wish to include edge attributes (e.g., information about the relationship between the two concepts) or node attributes (information about each concept). We show how to do so below. But first, consider the basic use of `review()`: 


```{r}
lit <- review(literature, from = "from", to = "to")

lit

head(lit$edgelist)

head(lit$nodelist)
```

Edge and node attributes can be added using the `edge_attributes` and `node_attributes` arguments. `edge_attributes` is a vector that identifies columns in the supplied data frame that the user would like to retain. `node_attributes` is a separate dataframe that contains attributes for each node in the primary data set. The example `node_attributes` data include one column `type` indicating a type for each each node/variable/concept. 

```{r}
data("node_attributes")

head(node_attributes)

lit <- review(literature,
              edge_attributes = c("cites", "cites_empirical"),
              node_attributes = node_attributes)

lit

head(lit$edgelist)

head(lit$nodelist)
```

Tip: to retain all variables from `literature`, use `edge_attributes = names(literature)`.


## More Advanced Uses: larger networks, visualizating your network, network descriptives

Additional columns in the redistricting literature data include discriptions of the `edge` (the relationship between the `to` and `from` concepts), the theorized `mechanism`, and `cite_weight`---the number of studies in the literature that cite that that causal relationship.  

### A Larger Edgelist and Nodelist

```{r}
# load expanded dataset
literature <- read_csv(here::here("data", "literature.csv")) 

# clean up text for better visualization 
literature %<>% 
  mutate(cites = cites %>% str_remove(",.*|;.*"))

literature$from %<>% str_replace_all("([a-z| |-]{8}) ", 
                           "\\1\n") 

literature$to %<>% str_replace_all("([a-z| |-]{8}) ", 
                           "\\1\n") 

literature$from %<>% str_replace_all(" ([a-z| |-]{9})", 
                           "\n\\1") %>% str_to_title()

literature$to %<>% str_replace_all(" ([a-z| |-]{9})", 
                           "\n\\1") %>% str_to_title()

literature$from %<>% str_replace("\nOf\n", "\nOf ")

literature$to %<>% str_replace("\nOf\n", "\nOf ")

literature$from %<>% str_replace("\nFellow ", " Fellow\n")

literature$to %<>% str_replace("\nFellow ", " Fellow\n")

literature$from %<>% str_replace("\nState\n", " State\n")

literature$to %<>% str_replace("\nState\n", " State\n")

literature$from %<>% str_replace("\nDistrict\n", " District\n")

literature$to %<>% str_replace("\nDistrict\n", " District\n")

literature$from %<>% str_replace("\nWith\n", " With\n")

literature$to %<>% str_replace("\nWith\n", " With\n")


literature$cite_weight %<>%  replace_na(0)
```


Create variables to identify subgraphs
```{r}
literature %<>% mutate(partisan = str_c(to, from) %>% str_detect("Partisan"),
                       comm = str_c(to, from) %>% str_detect("Commun"),
                       confound = case_when(
      from == "Preserve\nCommunities\nOf Interest" & to == "Rolloff" ~ T,
      from == "Voter\nInformation\nAbout Their\nDistrict" & to == "Rolloff" ~ T,
      from == "Preserve\nCommunities\nOf Interest" 
            & to == "Voter\nInformation\nAbout Their\nDistrict" ~ T,
      T ~ F),
              empirical = ifelse(!is.na(cites_empirical),
                                 "Empirical work", 
                                 "No empirical work"))

# now with all node and edge attributes 
lit <- review(literature,
              edge_attributes = names(literature),
              node_attributes = node_attributes
              )

edges <- lit$edgelist
head(edges)

nodes <- lit$nodelist
head(nodes)
```


## With `visnetowrk`
```{r visnetwork, fig.width=7, fig.height=6}

library(visNetwork)


# define function to plot
literature_plot <- function(lit){

  nodes <- lit$nodelist %>%
    mutate(id = node) %>% 
  filter(!is.na(id)) %>%
  distinct() %>%
    # removed nodes with multiple types
  add_count(id) %>% 
  filter(n == 1)

edges <- lit$edgelist %>% transmute(
  from = from,
  to = to,
  detail = paste(edge, mechanism, cites, sep = "<br>") %>% str_remove_all("NA"), 
  type = edge
) %>% 
  filter(!is.na(from),!is.na(to)) %>% 
  distinct()



# use betweeness to scale nodes
nodes$icon.size <-nodes$betweenness

# add attributes
nodes <- nodes %>% mutate(label = id, 
                          title = paste0("<p>", type, ": ", label,"</p>"),
                          # levels in case we want Hierarchical Layout
                          level = ifelse(type == "goal", 1:2, 3:4),
                          # FontAwesome.com shapes for fun
                          shape = "icon",
       icon.color = case_when(type =="goal" ~ "black",
                              type !="goal" ~ "black"),
       icon.code = case_when(type == "condition" ~ "f205", # chess board
                             type == "goal" ~ "f24e", # scale  "f05b", # crosshairs
                             type == "policy" ~ "f0e3", # gavel
                             type == "value" ~ "f004", # "f4be", # hand with heart
                             type == "effect" ~ "f080", # "f681", # data 
                             type == "metric" ~ "f1de",# "f548", # ruler 
                             TRUE ~ "f0c8"), #square
       icon.face =  "'FontAwesome'",
       icon.weight = "bold")

# format edges
edges <- edges %>% mutate(
  title = paste0("<p>", detail, "</p>"),
  #label = type,
  color = ifelse(str_detect(type, "^increase"), "#81a275", "#617d9f"),
  color = ifelse(str_detect(type, "^decrease"), "#b14552", color) ) 



# make directed graph
visNetwork(nodes=nodes, edges=edges, width = "100%") %>% 
  visEdges(width=5, 
           color= edges$color, 
           arrows = "to", 
           arrowStrikethrough = F, 
           smooth = T) %>%
  visNodes(scaling=list(min=40, max=50)) %>%
  visOptions(highlightNearest = list(enabled = T, 
                                     degree = 1, 
                                     hover = T)) %>%
  visInteraction(hover=TRUE, zoomView = TRUE) %>%
  #visHierarchicalLayout() %>% 
  visPhysics(solver = "forceAtlas2Based", 
             forceAtlas2Based = list(gravitationalConstant = -50)) %>%
  addFontAwesome(name = "font-awesome-visNetwork") %>%
visLayout(randomSeed = 12) # to have always the same network  
}
```

### The core nodes

```{r dag-core, fig.width=7, fig.height=6}
core <- literature %>% filter(core)

core %<>% review(
  edge_attributes = names(core),
  node_attributes = node_attributes)

# plot core nodes
literature_plot(core)
```

### The whole mess
```{r dag-whole, fig.width=7, fig.height=6}
# plot core nodes
literature_plot(lit)
```







