---
title: "Simulating Biased Literature Reviews with `netlit`"
author: "Devin Judge-Lord, Adeline Lo & Kyler Hudson"
subtitle: Redistricting Literature
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(knitr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = FALSE,
                      fig.width=10, 
                      fig.height=7,
                      fig.show="hold", 
                      out.width = "100%",
                      #split = T,
                      fig.align = 'center', 
                      fig.path='../figs/',
                      fig.cap= "",
                      out.extra = "",
                      fig.retina = 3, 
                      warning=FALSE, 
                      message=FALSE)

library(kableExtra) # Table formatting

# format kable for document type 
kable <- function(...){
  if(knitr::is_latex_output()){
    head(..., 25) %>% 
    knitr::kable(booktabs = T, format = 'latex') %>% 
  kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
    } else {
    knitr::kable(...) %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
    }
}
```


# Metadata 

## Including Author Gender, H-Index, and Citation

```{r literature}
library(netlit)
library(tidyverse)
library(magrittr)

# Load main data and metadata on citations
#FIXME data() WILL WORK WHEN NETLIT"S literature_metadata IS UPDATED 
# data("literature_metadata")
# load("data/literature.rda")
load("data/literature_metadata.rda")
# data("literature")
load("data/literature.rda")
#FIXME end

names(literature_metadata) %<>% janitor::make_clean_names()

literature_metadata %<>% 
  rename(author_gender = author_sex)

literature_metadata%>% kable()

# split out multiple cites per edge 
literature_long <- literature %>% 
  mutate(id = str_split(cites, ";")) %>% 
  unnest(id)

# merge edgelist with metadata
literature_long %<>% full_join(literature_metadata)
```

```{r bias-hist, fig.width=3.5, fig.height=2, out.width="50%", fig.show='hold'}
literature_long %>% 
  ggplot() +
  aes(x = author_h_index, fill = author_gender)+
  geom_histogram()

literature_long %>% 
  ggplot() +
  aes(x = author_citations, fill = author_gender)+
  geom_histogram()
```

---

## Missing metadata 
```{r}
literature_metadata %>%
  filter(is.na(author_gender) | is.na(author_h_index)| is.na(author_citations)) %>% 
  kable()
```


```{r libraries, message=FALSE, warning=FALSE}
library(ggraph)
```

```{r include=FALSE}
clean <- . %>% 
  str_replace_all("([a-z| |-]{8}) ","\\1\n") %>%
  str_replace_all(" ([a-z| |-]{9})",  "\n\\1") %>% str_to_title() %>% 
  str_replace("\nOf\n", "\nOf ") %>% 
  str_replace("\nFellow ", " Fellow\n") %>% 
  str_replace("\nState\n", " State\n") %>% 
  str_replace("\nDistrict\n", " District\n") %>% 
  str_replace("\nWith\n", " With\n")

literature_long$from %<>% clean() 
literature_long$to %<>% clean()
```

---


## The Full Graph

```{r ggraph, cache= FALSE}
lit <- literature_long %>% 
  distinct(to, from) %>% 
  review()

lit

# best seed 1,4, *5*
set.seed(5)

netlit_plot <- function(g){
ggraph(g, layout = 'fr') + 
  geom_node_point(
    aes(color = degree_total %>% as.factor() ),
    size = 6, 
    alpha = .7
    ) + 
  geom_edge_arc2(
    aes(
      start_cap = circle(3, 'mm'),
      end_cap = circle(6, 'mm'),
      color = edge_betweenness,
      ),
    curvature = 0,
    arrow = arrow(length = unit(2, 'mm'), 
                  type = "open")
    ) +
  geom_edge_loop(
      aes( color = edge_betweenness,
      start_cap = circle(5, 'mm'),
      end_cap = circle(2, 'mm'),
      ),
      n = 300,
      strength = .6,
    arrow = arrow(length = unit(2, 'mm'), 
                  type = "open")
    ) +
  geom_node_text( aes(label = name), size = 2.3) + 
  ggplot2::theme_void() + 
  theme(legend.position="bottom") + 
  labs(edge_color = "Edge Betweenness",
       color = "Total Degree\nCentrality",
       edge_linetype = "") + 
scale_edge_color_viridis(discrete = FALSE,
                         option = "plasma", 
                         begin = 0, 
                         end = .9, 
                         direction = -1) +
  scale_color_viridis_d(option = "mako", 
                        begin = 1, 
                        end = .5)
}


literature_long %>% 
  distinct(to, from) %>% 
  review()  %>% 
  .$graph %>% 
  netlit_plot()
```

---

## Men only

- 69 edges to 61 edges
- 56 nodes to 52 nodes


```{r}
literature_long %<>%
  mutate(author_is_man = author_gender == "M")

bro_lit <- literature_long %>% 
  filter(author_is_man) %>% 
  distinct(to, from) %>% 
  review()

bro_lit


literature_long %>% 
  filter(author_is_man) %>% 
  distinct(to, from) %>% 
  review()  %>% 
  .$graph %>% 
  netlit_plot()
```

Nodes missing: `r lit$nodelist$node[!lit$nodelist$node %in% bro_lit$nodelist$node]`

---

## Authors with H-Index over 20

- 69 edges to 56 edges
- 56 nodes to 49 nodes


```{r}
lit_sr <- literature_long %>% 
  filter(author_h_index >2) %>% 
  distinct(to, from) %>% 
  review()

lit_sr 

literature_long %>% 
  filter(author_h_index >2) %>% 
  distinct(to, from) %>% 
  review()  %>% 
  .$graph %>% 
  netlit_plot()
```

Nodes missing: `r lit$nodelist$node[!lit$nodelist$node %in% lit_sr$nodelist$node]`

---

## Authors with over 1000 Citations

- 69 edges to 26 edges
- 56 nodes to 32 nodes


```{r, }
lit_sr_cites <- literature_long %>% 
  filter(author_citations > 1000) %>% 
  distinct(to, from) %>% 
  review()

lit_sr_cites

literature_long %>% 
  filter(author_citations > 1000) %>% 
  distinct(to, from) %>% 
  review()  %>% 
  .$graph %>% 
  netlit_plot()
```

Nodes missing: `r lit$nodelist$node[!lit$nodelist$node %in% lit_sr_cites$nodelist$node]`

---

# Biased Samples

```{r}
# biased sample weights 
literature_long %<>% 
    mutate(unbiased = .5,
           weight = case_when(
      author_is_man ~ .6,
      !author_is_man ~ .4,
      TRUE~ .5 
    ))


# a function to sample the network 
sample_lit <- function(n, literature_long, prob){
  
  # create index for sample
  samp_idx <- sample(seq_len(nrow(literature_long)), 
                     100, # 100 draws = number of studies to draw 
                     prob=prob # with prob var provided 
                     )
  
  # subset sample to index 
  sample <- literature_long %>% 
    rowid_to_column() %>% 
    filter(rowid %in% samp_idx) %>% 
    distinct(to, from) %>% 
    review()
  
    return(sample)
}

# TEST 
# random_samples <- sample_lit(literature_long, prob =  literature_long$unbiased)
# gender_samples <- sample_lit(literature_long, prob = literature_long$weight)
```

---

## Random draws of 100 studies

```{r draw-random, out.width= "10%", cache=TRUE}
random_samples <- map(1:100, # 100 samples 
                      sample_lit,
                      literature_long=literature_long, 
                      prob = literature_long$unbiased)

# make table of total number of nodes, edges and the graph object for plotting
random <- tibble(
  nodes = random_samples %>% map(1) %>% modify(nrow) %>% unlist(),
  edges = random_samples %>% map(2) %>% modify(nrow) %>% unlist(),
  graph = random_samples %>% map(3),
  sample = "Random"
)

# map(random$graph, netlit_plot)
```


Average nodes recovered: `r random$nodes %>% mean() `

Average edges recovered: `r random$edges %>% mean() `

---

## Gender-biased draws 

### pr(cite|man) = .60, pr(cite|woman) = .40

```{r draw-gender-bias, out.width= "10%", cache=TRUE}
# gender biased samples
gender_samples <- map(1:100, sample_lit,literature_long=literature_long, prob = literature_long$weight)

gender <- tibble(
  nodes = gender_samples %>% map(1) %>% modify(nrow) %>% unlist(),
  edges = gender_samples %>% map(2) %>% modify(nrow) %>% unlist(),
  graph = gender_samples %>% map(3),
  sample = "Gender bias favoring men"
)
  
#map(gender$graph, netlit_plot)
```

Average nodes recovered: `r gender$nodes %>% mean() `

Average edges recovered: `r gender$edges %>% mean() `


---

### pr(man) = .100, pr(woman) = .30

```{r draw-gender-bias-extreme, out.width= "10%", cache=TRUE}
# biased sample weights 
literature_long %<>% 
    mutate(weight = case_when(
      author_is_man ~ 1,
      !author_is_man ~ .3,
      TRUE~ .5 
    ))


# gender biased samples
gender_samples <- map(1:100, sample_lit,literature_long=literature_long, prob = literature_long$weight)

gender <- tibble(
  nodes = gender_samples %>% map(1) %>% modify(nrow) %>% unlist(),
  edges = gender_samples %>% map(2) %>% modify(nrow) %>% unlist(),
  graph = gender_samples %>% map(3),
  sample = "Gender bias favoring men"
)
  
#map(gender$graph, netlit_plot)
```

Average nodes recovered: `r gender$nodes %>% mean() `

Average edges recovered: `r gender$edges %>% mean() `


---

### pr(man) = .30, pr(woman) = .100

```{r draw-gender-affirm, out.width= "10%", cache=TRUE}
# biased sample weights 
literature_long %<>% 
    mutate(weight = case_when(
      author_is_man ~ .3,
      !author_is_man ~ 1,
      TRUE~ .5 
    ))


# gender biased samples
gender_samples2 <- map(1:100, sample_lit,literature_long=literature_long, prob = literature_long$weight)

gender2 <- tibble(
  nodes = gender_samples %>% map(1) %>% modify(nrow) %>% unlist(),
  edges = gender_samples %>% map(2) %>% modify(nrow) %>% unlist(),
  graph = gender_samples %>% map(3),
  sample = "Gender bias favoring women"
)
  
#map(gender$graph, netlit_plot)
```

Average nodes recovered: `r gender$nodes %>% mean() `

Average edges recovered: `r gender$edges %>% mean() `


---

## H-Index-biased draws

(replacing NA HIndex with 0) 

```{r draw-hindex-bias, out.width= "10%", cache=TRUE}
literature_long %<>%
  mutate(author_h_index = replace_na(author_h_index, 0 ))

# gender biased samples
hindex_samples <- map(1:100, sample_lit,literature_long=literature_long, prob = literature_long$author_h_index)

hindex <- tibble(
  nodes = hindex_samples %>% map(1) %>% modify(nrow) %>% unlist(),
  edges = hindex_samples %>% map(2) %>% modify(nrow) %>% unlist(),
  graph = hindex_samples %>% map(3),
  sample = "H-Index bias"
)
  
#map(gender$graph, netlit_plot)
```

Average nodes recovered: `r hindex$nodes %>% mean() `

Average edges recovered: `r hindex$edges %>% mean() `

---

## Citation-biased draws

```{r draw-citation-bias, out.width= "20%", cache=TRUE, fig.show="hide"}
literature_long %<>%
  mutate(author_citations = replace_na(author_citations, 0 ))

# gender biased samples
citations_samples <- map(1:100, sample_lit,literature_long=literature_long, prob = literature_long$author_citations)

citations <- tibble(
  nodes = citations_samples %>% map(1) %>% modify(nrow) %>% unlist(),
  edges = citations_samples %>% map(2) %>% modify(nrow) %>% unlist(),
  graph = citations_samples %>% map(3),
  sample = "Citations bias"
)
  
# map(citations$graph, netlit_plot)
```

Average nodes recovered: `r citations$nodes %>% mean() `

Average edges recovered: `r citations$edges %>% mean() `


# Comparing biases

```{r biases, fig.width=4.5, fig.height=2, out.width="49%", fig.show="hold"}
s <- full_join(random, gender) %>% 
  full_join(hindex) %>% 
  full_join(citations)

s %>% 
  ggplot() + 
  aes(x = nodes, fill = sample, color = sample) +
  geom_density(alpha = .5) + 
  scale_color_viridis_d() + 
  scale_fill_viridis_d() +
  theme_minimal() + 
  labs(color = "", 
       fill = "") + 
  theme(axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

s %>% 
  ggplot() + 
  aes(x = edges, fill = sample, color = sample) +
  geom_density(alpha = .5) + 
  scale_color_viridis_d() + 
  scale_fill_viridis_d() +
  theme_minimal() + 
  labs(color = "", 
       fill = "") + 
  theme(axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
  
```
